---
title: "Exploratory data analysis of UFO data set"
author: "Anita Li, Jacob McFarlane, Steffen Pentelow, Chirag Rank"
date: "20/11/2020"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(feather)
library(ggplot2)
library(knitr)
```

# Summary of the data set


``` {r load data}
ufos <- read_feather("../data/raw/aliens.feather")

# basic cleaning
ufo_tidy <- ufos %>%
  mutate(approximate = if_else(str_detect(Duration, "~|About|<|>"),1,0),
         range = if_else(str_detect(Duration, "-|:|to"),1,0),
         decimal = if_else(str_detect(Duration, "\\."),1,0),
         nightly = if_else(str_detect(Duration, "night"),1,0),
         minute = if_else(str_detect(Duration, "((M|m)in)|MIN"),1,0),
         second = if_else(str_detect(Duration, "((S|s)ec)|SEC"),1,0),
         hour = if_else(str_detect(Duration, "((H|h)our)|HOUR"),1,0)) %>%
  filter(approximate == 0 & range == 0 & decimal == 0, nightly == 0) %>%
  mutate(numeric_vals = str_extract_all(Duration, "[:digit:]+")) %>%
  unnest(numeric_vals) %>%
  mutate(numeric_vals = as.numeric(numeric_vals), 
         clean_vals = case_when(
           hour == 1 ~ (numeric_vals * 3600),
           minute == 1 ~ (numeric_vals * 60),
           second == 1 ~ numeric_vals)) %>%
  drop_na(clean_vals, Shape) %>%
  filter(!Shape %in% c('Unknown', 'Other', 'Changing', '')) %>%
  select(`Date / Time`, City, State, Shape, clean_vals) %>%
  rename(
    data_time = `Date / Time`,
    duration_sec = clean_vals
  )

```


The data set used in this project are records of UFO sightings in [British Columbia](http://www.nuforc.org/webreports/ndxlBC.html), Canada and [Washington States](http://www.nuforc.org/webreports/ndxlWA.html), USA, which is provided by [America's foremost UFO Reporting Agency](www.ufocenter.com) since 1974. Each row in the data set represents an observation of UFO sighting, and features recorded include place and time, shape of UFO, duration of sightings, and a short descriptive summary. There are `r nrow(ufos)` observations and `r ncol(ufos)` features in the data set. However, there are many records with invalid shape and duration. After removing invalid records, there are `r nrow(ufo_tidy)` observations left. In this project, we only interest in UFO shapes that have more than 30 observations. Below is a summary of duration for each UFO shape.

```{r SUMMARY TABLE}
ufo_summary <- ufo_tidy %>%
  group_by(Shape) %>%
  summarise(median = median(duration_sec),
            count = n(),
            min = min(duration_sec),
            max = max(duration_sec)) %>%
  filter(count > 30) %>%
  arrange(median)

kable(ufo_summary, caption = "Table 1. Statistic summary on the duration (seconds) of sightings for each shape")

```




# Exploratory analysis on the data set

We choose median to represent the average duration for each sighting because the distribution of duration is skewed. To look at weather the median of duration are different for different shape, we plotted the distribution of duration (on log scale) for each shape. There are 14 shapes in this analysis, and for some of them, their medians seem to look the same (i.e. triangle and sphere).

```{r EDA, fig.width=10, fig.height=9, fig.cap="Distribution of time duration (in log scale) for each UFO shape"}
# get sorted order of shape
shape_order <- as.vector(ufo_summary$Shape)

ufo_tidy %>%
  filter(Shape %in% shape_order) %>%
  ggplot(aes(x = duration_sec, 
      y = factor(Shape, levels = shape_order))) +
  geom_jitter(height = 0.15, size = 2, alpha = 0.1) +
  stat_summary(data = ufo_summary,
               aes(x = median, y = Shape),
               geom = 'point', shape = 15,
               size = 2, color = 'red') +
  scale_x_log10(labels = scales::comma) +
  xlab("Duration (in log10 scale)") +
  ylab("UFO shape") +
  theme(legend.position = 'top',
        legend.title = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(size = 1),
        axis.ticks = element_line(size = 1),
        axis.ticks.length = unit(.25, 'cm'),
        text = element_text(size = 15),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), 'cm')
  )

```


